---
output: 
html_document:
  number_sections=TRUE
pagetitle: Assignment_4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# **Assignment 4**

In this assignment, you will use R (within R-Studio) to:

+ Use several methods to load external data files into R
+ Explore parameters available for customizing the way data is read from a file
+ Use base R tools to view attributes of data sets
+ Perform simple summary statistics
+ Use the built-in visualization methods to quickly look at relationships in your data
+ Demonstrate one method for saving visualizations to an image file

**All file paths should be relative, starting from *your* Assignment_4 directory!!**

**This means that you need to create a new R-Project named "Assignment_4.Rproj" in your Assignment_4 directory, and work from scripts within that.**


## **For credit...**

1.  Push a completed version of your Rproj and R-script (details at end of this assignment) to GitHub
2.  Your score will also depend on whether any files generated in this workflow are found in your repository
3.  There are 7 questions scattered through this walkthrough. Upload a plaintext file with numbered answers to the **bolded** assignment questions to Canvas. This shows that you worked through the assignment and lets me know to pull a fresh copy of your GitHub repo to grade.


____________


It would be terribly inconvenient if R made us manually enter our data. Thankfully, there are dozens (hundreds?)
of ways that we can read external data into R for analysis.  Most of the time the data we want to analyze comes
in the form of an Excel spreadsheet.  There are special ways to import Excel spreadsheets directly, but typically
we don't want to store our data as .xlsx because it's a large bloated file format. People who work with data in
"rectangular" format (like a spreadsheet) often use a form called "comma-separated-values," or .csv


We will use the built-in function read.table() to load some data.

```{r, echo=TRUE}
?read.table() #This brings up the help file
df <-  read.csv("../../Data/landdata-states.csv") # why did I change to read.csv ???
class(df) # what type of object is df?
head(df) # shows the first 6 elements of an object (first 6 rows if you give it a data frame)
```


Now, we have a data frame loaded into R and stored as an object called "df."
If you open that same file with a plain text editor you'd see a bunch of values separated by commas.
The read.csv() function is a convenient way to tell R that those commas separate different values and each "\\n" (newline) character means a new row. It automatically treats the first row as column headers.  

**Questions:**

  + **1. What other stuff does read.csv() do automatically?**
  + **2. How is it different from read.csv2()?**
  + **3. Why does read.csv2() even exist?**


Now, I notice that each column in this data frame has its own class. Let's look at a couple
```{r, echo=TRUE}
class(df$State)
class(df$Date)
```

**Questions:**

  + **4. How could I change the parameters of read.csv() to make it so the class of the "State" column is "character" instead of "factor?"**

____________


Now, let's explore this data set a bit with basic descriptive stats...

```{r, echo=TRUE}
dim(df) # dimensions of the data frame (rows, columns)
str(df) # another nice way to glimpse a data frame
summary(df) # summary() works differently for different data classes. Note how it summarizes factors vs numerics
```

**Questions:**

  + **5. What command would give the summary stats for ONLY the Home.Value column?**
  + **6. What value is returned by the command: names(df)[4]  ?**
  
____________


We can do some very basic visualizations of our data as well. In many cases, a good image is much more descriptive than a boring table of summary statistics...

```{r, echo=TRUE}
hist(df$Land.Value) # histogram showing number of times each numeric value was seen in the vector "Land.Value"

# If you want to look at land value by region, you could do this:
plot(x=df$region,y=df$Land.Value)

# Land value by year
plot(x=df$Year,y=df$Land.Value)
```

Note that the plot() function automatically tries to determine the best type of plot for your data based on the classes of vectors that you give it.

**Questions:**

  + **7. What is happening when you add (...col=df\$region) to the above plotting code?**  
  **In other words, what happens when you run: plot(x=df\$Year,y=df\$Land.Value,col=df\$region)**
  
_______________

# Now, for the rest of the assignment...

1.  Create a new R script as part of your Assignment 4 R-project. Name it "Assignment_4.R"
2.  That script should do the following:
    + Read in the file: "/Data/ITS_mapping.csv" (filepath starts in *your* Data_Course main directory)  ...good luck with that, hahaha!
    + Somehow summarize all of the columns and do a bit of additional exploration (play with some functions)
    + Make a boxplot where "Ecosystem" is on the x-axis and "Lat" is on the y-axis
    + Write code to export this boxplot to a new file in your Assignment_4 directory called "silly_boxplot.png"
      Hints on below ...
3.  Make sure to save your completed script and Rproject and make sure your png file is saved correctly
4.  Push all these saved changes and new files onto your GitHub repository so I can grade them
5.  Don't forget the plaintext file with answers to bolded questions needs to go to Canvas as well!      

_____________


To use the base R method to save a plot to an image file, you just wrap your code for the image between two commands.

```{r, eval=FALSE, include=TRUE}
png(filename = "./silly_boxplot.png")
#whatevercodeyoucameupwithforyourplot
dev.off()
```

The png() function has lots of options you can tweak, but it opens a "graphics device" that starts collecting any output from R into a .png image file.

The dev.off() function just closes the graphics device and writes the previous input to the file you specified.