---
output:
  html_document: default
  pdf_document: default
---
#**Data Analysis Course Outline**

###WEEK 1

* Why use code!?
* INSTALL R / R-studio
* Familiarize R-studio functions and layout
* Where to look for help

```{r eval=FALSE, include=TRUE}
# Functions covered (parital list):
help()
```
  
###WEEK 2

* Assigning things to objects
* Get familiar with object types and basic functions
    * values, vectors, lists
    * data frames and matrices
    * boolean, character, numeric, POSIXct  
* Accessing elements of objects
* Boolean evaluations
* Data-type conversions


```{r eval=FALSE, include=TRUE}
# Functions covered (parital list):
=       <-      ->
class()   
data.frame()  as.factor()   as.numeric()    as.character()    as.POSIXct()    as.matrix()
==    <   >   <=    >=
+     -   *   /     
which()   signif()    ceiling()     floor()       round()    
c()       list()      cbind()       rbind()       sum()         mean()        sd()        
```


  
###WEEK 3

* Importing data
* Useful data formats
* Data structure and attributes
* Summary stats and basic visualizations
* Exploring data
    * Sorting, Transposing, Sampling
    * heatmaps, boxplots, barcharts, scatterplots, histograms

```{r eval=FALSE, include=TRUE}
# Functions covered (parital list):
read.csv()        read.delim()
str()             dim()             names()           attributes()          head()          tail()
summary()         min()             max()             range()               quantile()      var()
hist()            boxplot()         barplot()         plot()                heatmap()
sample()          t()               sort()
```
  
###WEEK 4

* Finding/Installing/Loading packages
* Extending functionality of R
* Subsetting and manipulating raw data
* Output options

```{r eval=FALSE, include=TRUE}
# Functions covered (parital list):
install.packages()      
library()
cor()
write.table()       sink()
```
      Skills Test 1:

      * Import data set
      * Convert elements to new data type
      * Subset based on values
      * Calculate summary statistics
      * Create basic summary figures
      * Export summary statistics to text file
  

###WEEK 5

* Data estimations
    * point estimates
    * interval estimates
* Hypothesis testing / Model fitting
    * t-test (paired/unpaired)
    * chi-square
    * ANOVA
    * LM/GLM

```{r eval=FALSE, include=TRUE}
# Functions covered (parital list):
lm()        glm()         aov()       summary()
t.test()    chisq.test()  
```

###WEEK 6

Experimental design
common designs and analysis options
quantitative vs qualitative data
Probability distributions
Fitting distributions
Type I and Type II errors
Post-hoc tests

```{r eval=FALSE, include=TRUE}
# Packages used (partial list):
fitdistrplus
MASS
# Functions covered (parital list):
plotdist()      descdist()
fitdist()       denscomp()        cdfcomp()
TukeyHSD()      
```

###WEEK 7

* Non-parametric alternatives
* Mann-Whitney-Wilcoxin
* Kruskal-Wallace
* Apply functions

```{r eval=FALSE, include=TRUE}
# Packages used (partial list):

# Functions covered (parital list):
wilcox.test()     kruskal.test()

apply()     sapply()      lapply()      tapply()
```
###WEEK 8

* Other peoples' data
* Principles of tidy data
* Intuitive manipulations and group functions
    * filter
    * arrange
    * select
    * mutate
    * group_by
    * summarize
    * %>%
* Tidy data transformations
    * gather
    * spread

```{r eval=FALSE, include=TRUE}
# Packages used (partial list):
dplr      plyr      tidyr
# Functions covered (parital list):
filter()      arrange()       select()        mutate()        
group_by()    summarize()     %>%
gather()      spread()
```
      Skills Test 2: 	
      
      * Import messy data
      * Convert to tidy format
      * Plot data distribution
      * Rearrange and mutate data set
      * Summary stats on grouped data
      * Test hypothesis / post-hoc tests
      

###WEEK 9
* Predicting data
* Intro to ggplot	

```{r eval=FALSE, include=TRUE}
# Packages used (partial list):
ggplot2

# Functions covered (parital list):
predict() 
qplot()
ggplot()
    aes()
```
    
###WEEK 10
* Figure generation
* Figure export

```{r eval=FALSE, include=TRUE}
# Packages used (partial list):
ggplot2

# Functions covered (parital list):
ggplot()
geom_point()      geom_boxplot()      geom_bar()      geom_violin()
labs()            ggsave()            
```

###WEEK 11
* Figure generation continued

```{r eval=FALSE, include=TRUE}
# Packages used (partial list):
ggplot2
gridExtra

# Functions covered (parital list):
grid.arrange()
ggplot()
scale_*()
```

###WEEK 12
* Data standardization / normalization
* Ecology examples
    * Ordinations / NMDS
    * PermANOVA
    * Distance measures
    * Diversity measures

```{r eval=FALSE, include=TRUE}
# Packages used (partial list):
vegan

# Functions covered (parital list):
decostand()       rrarefy()         dist()        betadiver()         diversity()         betadisper()
metaMDS()         adonis()
```
   
   
      Skills Test 3:	
      
      * Import data set
      * Fit appropriate model
      * Use model to predict new response values from new predictors
      * Generate and export plots from data sets

###WEEK 13

* Importing and manipulating DNA sequence data
    * Bioconductor
    * Sequence data
    * Biostrings
* Phylogenetics examples
    * Sequence alignment
    * Tree building
    * Taxonomic assignment

```{r eval=FALSE, include=TRUE}
# Packages used (partial list):
Bioconductor
ape
biostrings

# Functions covered (parital list):


```

###WEEK 14

* Command-line tools
* BASH
* compression
* grep, sed, find, |


###WEEK 15
* Data management
* Reporting
* Rmd

