---
output:
  pdf_document: default
  html_document: default
---
### This will use the "tidyverse" suite of packages
```{r}
library(tidyverse)
```


## Exercise 1

Consider the following:
```{r,include=TRUE,eval=FALSE}
vector = "Good morning! "
```
**How many characters are in "vector" ?**


## Exercise 2

If:
```{r}
x <- c("Open", "Sesame ")
y <- c("You","Suck.")
nchar(x)
```

**Then, what is the value of:**
```{r,include=TRUE,eval=FALSE}
nchar(c(x,y))
```


## Exercise 3

If:
```{r}
m <- "The capital of the United States is Washington, D.C."
unlist(str_split(m," "))
```

...And:
```{r}
str_trunc(m,11, ellipsis = "")
```

...And:
```{r}
str_sub(m,start = 13,end = 25)
```


**Come up with a way to extract "Washington" from m.**


## Exercise 4

If:
```{r}
paste(m,", you idiot!", sep = "")
```

**Then come up with a way to use the vector "m" to paste together "United States, you idiot!"**


## Exercise 5

If:
```{r}
q = "What is the capital of the United States?"
c(q, paste0(m,", you idiot!"))
```

**Then, what will be the value of "d" for:**
```{r, include=TRUE}

d = str_split(c(q, paste0(m,", you idiot!")), pattern = " ")
```

## Exercise 6

If:
```{r}
c(unlist(map(d,1)),"Heck!?")
```

And:
```{r}
unlist(map(d,2))
```
**Then what does map(d,1) do?**  **...And why did I wrap it in unlist()**


## Exercise 7

If:
```{r}
t = c("a","ab","c","d","e","fa")
grep("a",t)
grepl("a",t)
```
...And:
```{r}
f = c("b","ca","at","c","e","aa")
v = list(f,t)
v
grep("a",v)
grepl("a",v)
```


**Then what will be the values of the following two expressions?**
```{r,include=TRUE,eval=FALSE}
grep("What",d)
grepl("What",d)
```


## Exercise 8

If:
```{r}
q
str_replace(q,"a","A")
```
**Then write some code to change ALL spaces to underscores "_" in the vector q**
