---
output:
  pdf_document: default
  html_document: default
---
## Exercise 1

If:
```{r}
X = c(NA,3,14,NA,33,17,NA,41)
is.na(X)
!is.na(X)
```

**Then what will be the value of:**
```{r,include=TRUE,eval=FALSE}
X[!is.na(X)]
```


## Exercise 2

If:
```{r}
Y = 21:28
Z=data.frame(X,Y)
Z
Z[is.na(Z)] <- 0
Z
```

**Then write an expression that will replace all instances of "NA" with the number 10 in the following vector "P":**

```{r}
P = c(X,33,NA,400,12,0,15)
P
```


## Exercise 3

Consider the following vector:
```{r}
W <- c (11, 3, 5, NA, 6)
```
**Write some R code that will return TRUE for the missing element in the vector W.**


## Exercise 4


Given the vector A:
```{r}
A <- c (33, 21, 12, NA, 7, 8)
```
**Write some code that will calculate the mean of A.**


## Exercise 5

Load ‘Orange’ dataset...
```{r}
data(Orange) 
head(Orange) # Take a look at first 6 rows
```
**Replace all values of age=118 with NA.**


## Exercise 6

For the following data frame X:
```{r}
c1 <- c(1,2,3,NA)
c2 <- c(2,4,6,89)
c3 <- c(45,NA,66,101)
X <- data.frame(c1,c2,c3) 
```

If:

```{r}
X
complete.cases(X)
```
**Write some code that will display only the rows with missing values.**


## Exercise 7

Given the following data frame:
```{r}
df <- data.frame (Name = c("NA", "Joseph", "Martin", NA, "Andrea"), 
                  Sales = c(15, 18, 21, 56, 60), 
                  Price = c(34, 52, 21, 44, 20), 
                  stringsAsFactors = FALSE) 
df
```
**Write some code to remove rows where "Name" is missing.**

