---
output:
  pdf_document: default
  html_document: default
---

### For the following data frame:
```{r}
data("mtcars")
head(mtcars)
```
# Keep in mind the following...
```{r}

x <- 1:10 # create vector
x >= 3 # greater than or equal to
x < 9 # less than 9
x > 3 & x < 7 # greater than 3 AND less than 7
x >= 9 | x <= 2 # greater than or equal to 9 OR less than or equal to 2
x %in% 5:10 # present in a given set
x == 3 # equal to 3
x != 3 # not equal to 3
x %% 2 == 0 # modulo division (does remainder equal 0 when divided by 2)
xor(x == 3, x > 1) # matches ONLY ONE of the conditions...NOT BOTH
```


## Exercise 1
Use logical operators to output only those rows of  mtcars  where column  mpg  is between 15 and 20 (excluding 15 and 20).

## Exercise 2
Use logical operators to output only those rows of  mtcars  where column  cyl  is equal to 6 and column  am  is not 0.

## Exercise 3
Use logical operators to output only those rows of  mtcars  where column  gear  or  carb  has the value 4.

## Exercise 4
Use logical operators to output only the even rows of  mtcars.

## Exercise 5
Use logical operators and change every fourth element in column  mpg  to 0.

## Exercise 6
Output only those rows of  mtcars  where columns  vs  and  am  have the same value 1, solve this without using  ==  operator.

## Exercise 7
(TRUE + TRUE) * FALSE , what does this expression evaluate to and why?

## Exercise 8
Output only those rows of  mtcars  where at least  vs  or  am  have the value 1, solve this without using  ==  or  !=.

## Exercise 9
Explain the difference between  | ,  || ,  &  and  &&.

## Exercise 10
Change all values that are 0 in the column  am  in  mtcars  to 2.

## Exercise 11
Add 2 to every element in the column  vs  without using numbers.

## Exercise 12
Output only those rows of  data  where  vs  and  am  have different values, solve this without using  ==  or  !=.