---
output:
  pdf_document: default
  html_document: default
---
### This will use the tidyverse package and the "gapminder" data set

```{r}
library(tidyverse)
library(gapminder)
data("gapminder")
df = gapminder
head(gapminder)
```
## Exercise 1

```{r}
levels(df$continent)
```
**Add "Antarctica" to the levels of possible continents**

## Exercise 2
**Add 3 more continents to the factor levels: "North America", "South America", and "Central America"**

## Exercise 3

**Consider the following groupings of countries**
```{r}
south <- c("Argentina","Bolivia","Brazil","Chile","Colombia","Ecuador",
           "Paraguay","Peru","Uruguay","Venezuela")
north <- c("United States","Canada","Mexico")
central <- c("Costa Rica","El Salvador","Guatemala","Honduras","Nicaragua","Panama")
```
**Currently, they are all in the "Americas" continent group.**
**Write some code to move them into their new respective continents**

## Exercise 4
**Clean up the "Continent" categories to remove "Americas" and "Antarctica"**
```{r}
levels(df$continent)
```

## Exercise 5
**Statistical models often use the FIRST level of a factor as a reference group**
**Reorder the levels of the "Continent" factor so that "North America" will be the reference group**


## Exercise 6

Ordered Factors:
```{r}
status <- c("Lo", "Hi", "Med", "Med", "Hi","Lo", "Hi", "Med", "Med", "Hi")
status <- factor(status)
ordered.status <- factor(status, levels = c("Lo","Med","Hi"), ordered = TRUE)

table(status)
table(ordered.status)
```
**Some functions in R treat ordered factors differently than unordered factors**
**What is different about an ordered factor?**



## Exercise 7

```{r}
as.numeric(status) + 1
```
**Figure out what is going on with the expression above.**


## Exercise 8


```{r}
sub_status <- status[status %in% c("Lo","Hi")] # extract only the "Lo" and "Hi" values
sub_status
```
**Re-factor "sub_status" to remove the "Med" level**


## Exercise 9

```{r}
status2 <- factor(c("Med", "Hi","Lo", "Hi", "Medium High"))
c(status,status2)
```
**Find a way to properly combine the two factored vectors "status" and "status2"**
**Make sure the resulting vector is a factor as well**

## Exercise 10

Look at the following plot:
```{r,fig.height=3,fig.width=3}
ggplot(df,aes(x=continent,y=lifeExp)) + geom_boxplot()
```
**Re-make the plot so that the bars are in descending order**


## Exercise 11

#### If:
```{r,fig.height=3,fig.width=3}
ggplot(df, aes(x=year,y=pop)) + geom_point()
```

#### And ...
```{r,fig.height=3,fig.width=3}
df$PopFactor <- cut(df$pop,5,labels = c("Really Low","Low","Med","High","Really High"))
ggplot(df, aes(x=year,y=PopFactor)) + geom_point()
```

**Convert the gdpPercap to an ordered factor with 6 levels and plot it against year**