---
output:
  pdf_document: default
  html_document: default
---
#### Consider the following:
```{r}
for(i in 1:3){
  x <- paste0("Number ",i)
  print(x)
}
```

#### And...
```{r}
planets <- c("Mercury","Venus","Earth","Mars","Jupiter","Saturn","Uranus","Neptune")
n <- 1
newvector <- c()
for(i in planets){
  newvector[n] <- paste0(i,i)
  n=n+1
  } 
newvector


```


## Exercise 1

**Write a for loop that iterates over the numbers 1 to 7 and prints the cube of each number using print().**

## Exercise 2

**Write a for loop that iterates over the column names of the inbuilt iris dataset and print each together with the number of characters in the column name in parenthesis. Example output: Sepal.Length (12). Use the following functions print(), paste0() and nchar().**

## Exercise 3

```{r}
i <- 1
while (i < 6) {
print(i)
i = i+1
}
```


**Write a while loop that prints out standard random normal numbers (use rnorm()) but stops (breaks) if you get a number bigger than 1.**

## Exercise 4

**Using a for loop simulate the flip a coin twenty times, keeping track of the individual outcomes (1 = heads, 0 = tails) in a vector that you preallocte.**


## Exercise 5

**Use a while loop to investigate the number of terms required before the product 1*2*3*4... reaches above 10 million.**

## Exercise 6

**Use a while loop to simulate one stock price path starting at 100 and random normally distributed percentage jumps with mean 0 and standard deviation of 0.01 each period. How long does it take to reach above 150 or below 50?**
