
### The gapminder package has interesting chronological data of life expectancy, GDP, etc.

#### Let's look at it over time

```{r,echo=TRUE,warning=FALSE,message=FALSE}
library(tidyverse)
library(gganimate)
library(gapminder)
```

```{r}
names(gapminder)
p1 = ggplot(gapminder, aes(x=gdpPercap,y=lifeExp,color=continent,group=country)) +
  geom_point() 
p1
```

#### Let's change that x-axis scale

```{r}
p1 + scale_x_log10()
```


Let's make point size reflect country population...

```{r}
p2 = ggplot(gapminder, aes(x=gdpPercap,y=lifeExp,color=continent)) +
  geom_point(aes(size=pop),alpha=.5) +
  scale_x_log10() + geom_smooth(se=FALSE)
p2
```

#### But what about the "time" component?
##### Can we add "year" as color maybe?

```{r}
ggplot(gapminder, aes(x=gdpPercap,y=lifeExp,color=year)) +
  geom_point(alpha=.2) + scale_x_log10() +
  geom_smooth(se=FALSE,method="lm",aes(group=year))

```

### Animations can help us get around trying to show too many things at once.

#### They aren't always a great idea, but can be handy.

```{r,echo=FALSE,warning=FALSE,message=FALSE}

ggplot(gapminder, aes(x=gdpPercap, y=lifeExp, size = pop, colour = country)) +
  geom_point(alpha = 0.7, show.legend = FALSE) +
  scale_colour_manual(values = country_colors) +
  scale_size(range = c(2, 12)) +
  scale_x_log10() +
  facet_wrap(~continent) + labs(subtitle = "Points scaled to country population") 
  # Here comes the gganimate specific bits
  labs(title = 'Year: {frame_time}', x = 'GDP per capita', y = 'life expectancy') +
  transition_time(year) +
  ease_aes('linear')
```