---
title: "Mushroom Growth"
author: "Zahn"
date: "3/12/2020"
output: html_document
---


### This experiement was one that I did during grad school when I was growing shrooms for the Farmers' Market.


First step is to load data (code not shown)
```{r, echo=FALSE,message=FALSE,warning=FALSE}
library(tidyverse)
dat <- read_csv("../Data/mushroom_growth.csv")

```

Quick look at our data set:
```{r}
glimpse(dat)
```


### Here's a plot showing Nitrogen effect on two species of oyster mushrooms:
```{r,echo = FALSE}
ggplot(dat, aes(x=Nitrogen,y=GrowthRate,color=Species)) +
  geom_point() + geom_smooth(method = "lm", formula = y ~ poly(x,3)) + theme_bw()

```

#### I'd like to model this, but include Light and Humidity, etc.

```{r, echo=FALSE,warning=FALSE,message=FALSE}
mod1 <- glm(data = dat, GrowthRate ~ (poly(Nitrogen,3) * Species) + Light + Humidity + Temperature)

MASS::stepAIC(mod1)

goodmodel <- glm(GrowthRate ~ poly(Nitrogen, 3) + Species + Light + 
    Humidity + Temperature, data = dat)


```


#### Our best model wound up being:
GrowthRate ~ poly(Nitrogen, 3) + Species + Light + 
    Humidity + Temperature
    
```{r}
summary(goodmodel)
```


#### Which combo of factors leads to highest growth rate???


```{r}

ggplot(dat,aes(x=Humidity,y=GrowthRate,fill=Species)) + geom_violin()
```

### For future projects, use *P. cornucopiae* at high humidity for best chance of fastest growth!
___

The end
