---
output: html_document
---

### Working with files and paths in R

Usually, the data you analyze will be sitting in an excel or csv file. You will have to find it and import it into R using code. Also, you will probably need to output things from R like statistical tables, graphs, or cleaned data into new files.  All of these tasks require you to be able to navigate your computer storage. Let's take a look at a few things we can do within R:

This command shows us the "path" to our current working directory. A path is like a set of directions for how to get from the very root of your computer to the current folder you're working in
```{r}
getwd()
```
"/" represents the *root* of my entire filesystem, and each slash represents a new subdirectory.


You can list the files in my current working directory with the following:
```{r}
list.files()[1:10] # just the first 10 to save space
```
That command can be modified as well:
```{r}
list.files(pattern = "x") # just filenames that have "x" in them
```
You can search within any directory on your computer, by telling list.files() which "path" to search in:
```{r}
list.files(path = "/home/gzahn/Desktop/Bioinformatics/")
```

```{r}
list.files(path = "/home/gzahn/Desktop/Bioinformatics/",
           recursive = TRUE,
           pattern = ".nex")

```
Note how "recursive = TRUE" tells it to descend into subdirectories of a given path. Those 4 files live in the "Fungal_Alignments" subdirectory within that path.

Now a closely related function:
```{r}
mypath <- "~/Desktop/GIT_REPOSITORIES/Data_Course/Data"
list.dirs(path = mypath, recursive = FALSE)
```

You can save this list of directories in case you want to work with it later:
```{r}
data_directories <- list.dirs(path = mypath, recursive = FALSE)
data_directories[3]
list.files(path = data_directories[3],full.names = TRUE)
```

You can ask questions about whether files or directories exist in a given location:
```{r}
file.exists("/home/gzahn/Desktop/GIT_REPOSITORIES/Data_Course/Data/flights/2679884.csv")
dir.exists("/home/gzahn/Desktop/GIT_REPOSITORIES/Data_Course/Data/fights") # misspelled "flights"
```
You can create and modify and peek inside files as well:
```{r}
list.files(path = data_directories[3],full.names = TRUE)
file.create(file.path(data_directories[3],"testfile")) # Says "TRUE" if it worked
list.files(path = data_directories[3],full.names = TRUE)

# be careful using file.remove() ... it's permanent!
file.remove("/home/gzahn/Desktop/GIT_REPOSITORIES/Data_Course/Data/flights/testfile") # Says "TRUE" if it worked
```

Here are some other functions you should play with:
```{r eval=FALSE}
file.rename()
file.append()
file.copy()
file.size()
readLines()
```


## The most important thing is to make sure you know how file paths *work!*

Do you know what is going on with the next 4 lines of code?
```{r}
getwd()
list.files()[1:10]
list.files(path = "..",full.names = TRUE)
list.files(path = "../Assignments")
```
___

This is all VERY useful once you start working with hundreds or thousands of data files for a given project. If I want to search my entire computer desktop and all the folders inside of it for fasta DNA sequence files and find the ones that match a pattern in naming:
```{r warning=FALSE}
mypath <- "~/Desktop"
fastas <- list.files(mypath,recursive = TRUE,pattern = "*5.8S*.fasta$",full.names = TRUE) # any file that has "5.8S" in the name and ends with ".fasta"
fastas
```
Since R did all the searching and saved the location of those files, I can have it automatically read them in and work with them. For example:
```{r}
fna <- ShortRead::readFasta(fastas)
ShortRead::sread(fna)
```

#### What is the absolute path to your Desktop on your computer?

#### Can you list all the files there?

#### Can you navigate around and find files in a different directory?
