---
title: "Untitled"
author: "gzahn"
date: "March 14, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=FALSE)
```

# This is an awesome way to make plots.

```{r, echo=FALSE}
library(tidyverse)

# Data
a <- data.frame( x=rnorm(20000, 10, 1.9), y=rnorm(20000, 10, 1.2) )
b <- data.frame( x=rnorm(20000, 14.5, 1.9), y=rnorm(20000, 14.5, 1.9) )
c <- data.frame( x=rnorm(20000, 9.5, 1.9), y=rnorm(20000, 15.5, 1.9) )
data <- rbind(a,b,c)
```

### Basic scatterplot
```{r, echo=FALSE}
ggplot(data, aes(x=x, y=y) ) +
  geom_point(alpha=.25)
```

### Density plot, instead
```{r,echo=FALSE}
ggplot(data, aes(x=x, y=y) ) +
  geom_bin2d() +
  theme_bw()
```

### Correlation plot
This is a correlation plot, that looks at multiple variables at the same time. Making it easy to get started analyzing your data.

```{r, echo=FALSE}
# Data example
data=mtcars[ , c(1,3:6)]

#Make the plot
plot(data , pch=20 , cex=1.5 , col=rgb(0.5, 0.8, 0.9, 0.7))
```

### This is a fancier version using the "car" package.
It's *basically* the same thing, but let's us add color groupings automatically.

```{r,echo=FALSE,fig.width=50,warning=FALSE}
library(car)
library(RColorBrewer)

data=mtcars

# Make the plot
my_colors <- brewer.pal(nlevels(as.factor(data$cyl)), "Set2")
scatterplotMatrix(~mpg+disp+drat|cyl, data=data , reg.line="" , 
                  smoother="", col=my_colors , smoother.args=list(col="grey"), 
                  cex=1.5 , pch=c(15,16,17) , main="Scatter plot with Three Cylinder Options")
```




### Density plot...
... for when a histogram is too blah.

```{r,echo=FALSE,warning=FALSE}
ggplot(data, aes(x=mpg,fill=factor(am))) +
  geom_density(alpha=.5)
```


# Testing the fancy animations

```{r,echo=FALSE,warning=FALSE}
library(gganimate)
library(gapminder)
data("gapminder")
data = gapminder

p1 = ggplot(data,aes(x=year,y=lifeExp,color=continent,frame=year)) +
  geom_point() + geom_smooth()


ggplot(gapminder, aes(gdpPercap, lifeExp, size = pop, colour = country)) +
  geom_point(alpha = 0.7, show.legend = FALSE) +
  scale_colour_manual(values = country_colors) +
  scale_size(range = c(2, 12)) +
  scale_x_log10() +
  facet_wrap(~continent) +
  # Here comes the gganimate specific bits
  labs(title = 'Year: {frame_time}', x = 'GDP per capita', 
       y = 'life expectancy') +
  transition_time(year) +
  ease_aes('linear')
```





